Trends of SDR in Thessaloniki

### 2023-06-28 Article.Rmd
Commit 2023-06-28 20:17

output: 
  rticles::elsevier_article:
    keep_tex:         no
    citation_package: natbib
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
  bookdown::word_document2: default
# date:           "`r format(Sys.time(), '%F')`"
---

@@

Describe the variability of the phenomenn

the inpacts of the radiation

the as far studies/coauses

The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
Changes of the SDR can be related to major changes on the Earths energy budget, the climate change mechanisms and the water and carbon cycles [@Wild2009].
Studies of SDR variability, have identify some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
There is a general consensus that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.

- observed long term changes
  europe / asia africa regions

Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that .

>>>>

output: 
  rticles::elsevier_article:
    keep_tex:         no
    citation_package: natbib
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              no
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
  bookdown::word_document2: default
# date:           "`r format(Sys.time(), '%F')`"
---

@@

The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
<!-- Trends -->
@Yang2021 reports a general dimming before 2000 and a brightening afterwards on multiple station in China. 
@Wild2021 shows a similar pattern, with the breaking point around 1980 from a 71 years record of a Central Europe station and for Brazil by @Yamasoe2021.
<!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->
@Yuan2021 with AI aided spatial analysis on continent level, containing numerous observational station, reach similar conclusions for the above regions and the global trend.

Although, there is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.

Due to the variability of those factors and the interaction between them,

...... cloud formation occurrences and constitution nuclei. cite.....

it is difficult to estimate the SDR trends on each site.

Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
<!-- causes say less we don't do that -->
